---
main_topsize: 0.15 #percent coverage of the poster
main_bottomsize: 0.08
primary_colour: "#1d428a"
secondary_colour: "#3779fa"
accent_colour:
#ESSENTIALS
title: '**Data Visualizations for NBA Analytics**'
title_textsize: 85pt
author:
  - name: '**Jiang Wei(David) Liu**'
    main: true
    email: jw8liu@uwaterloo.ca
affiliation:
    address: STAT442 - Visualization, University of Waterloo
main_findings:
  - "**Visualization** over **2021-22 NBA season** with **Tidyverse** + **Huxable**."
logocenter_name: NBA-Logo.png
output: 
  posterdown::posterdown_betterport:
    self_contained: TRUE
    pandoc_args: --mathjax
    number_sections: false

knit: pagedown::chrome_print
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html")
```


```{r, include=FALSE}
knitr::write_bib(c('posterdown', 'rmarkdown','pagedown'), 'packages.bib')
```

```{r, echo=FALSE, warning=FALSE}
library(tidyverse)
library(png)
library(ggpubr)
library(grid)
library(fmsb)
library(huxtable)
library(RCurl)
```

```{css, echo=FALSE}
.mybreak {
   break-before: column;
}
```

```{r, warning=FALSE}
# Import dataset and image
pbp <- read_csv("NBA_Playbyplay_2021-22.csv", show_col_types = FALSE)
pbox <- read.csv("NBA_Player_Boxscore_2021-22.csv")
tbox <- read.csv("NBA_Team_Boxscore_2021-22.csv")
bball_court <- readPNG("bball_court.png")
```

# Shot Attempts Heat Map

## 2D Continuous Element

The new generation players like Trae Young are shooting more **three-pointers** than ever, while the last **mid-range** master DeMar DeRozan scores most of his points inside the three point line.

```{r, echo=FALSE, warning=FALSE}
# Data Cleaning
trae <- pbp %>%
  filter(participants_0_athlete_id == 4277905) %>%
  filter(shooting_play == TRUE) %>%
  filter(coordinate_x > 0)

# Get Trae headshot pic
trae_pic <- readPNG(getURLContent(unique(filter(pbox, athlete_id == 4277905)$athlete_headshot_href)))

# Graph the heat map plot
trae_hm <- ggplot(data = trae, aes(coordinate_x, coordinate_y)) +
  labs(x="Baseline (ft)", y="Sideline (ft)",
       title = "Trae Yound (Ice Trae)",
       subtitle = "2021-22 Field Goal Attemps Heatmap") +
  # Add a basketabll court as the background image
  background_image(bball_court) +
  # Geom_count is Geom_point but size changes with frequency
  geom_count(color = "blue", alpha=0.6) +
  #adjust the scales
  scale_size_continuous(range = c(0.1, 4),
                        breaks = c(1,10,30),
                        name = "Number of Attempts",
                        ) +
  
  # Add a transparent layer of contour plot
  stat_density2d_filled(alpha = 0.45, 
                        contour_var = "count",
                        breaks = c(0,0.035,0.5,1.5,3,6,10),
                        show.legend = FALSE) +
  scale_fill_manual(values = c("#ffffff", "#cce1ff", "#99c3ff", 
                               "#66a6ff", "#3388ff", "#006aff")) +
  scale_x_continuous(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0,60)) +
  # make x and y 1-to-1
  coord_equal() +
  # Add a Trae headshot pic on the top left corner
  annotation_custom(rasterGrob(trae_pic), xmin = 0, xmax = 20, 
                    ymin = 42, ymax = 62)

trae_hm
```

```{r, echo=FALSE, warning=FALSE}
# Data Cleaning
trae <- pbp %>%
  filter(participants_0_athlete_id == 3978) %>%
  filter(shooting_play == TRUE) %>%
  filter(coordinate_x > 0)

# Get Trae headshot pic
trae_pic <- readPNG(getURLContent(unique(filter(pbox, athlete_id == 3978)$athlete_headshot_href)))

# Graph the plot
trae_hm <- ggplot(data = trae, aes(coordinate_x, coordinate_y)) +
  labs(x="Baseline (ft)", y="Sideline (ft)",
       title = "DeMar DeRozan",
       subtitle = "2021-22 Field Goal Attemps Heatmap") +
  # Add a basketabll court as the background image
  background_image(bball_court) +
  # Geom_count is Geom_point but size changes with frequency
  geom_count(color = "red", alpha=0.6) +
  #adjust the scales
  scale_size_continuous(range = c(0.1, 4),
                        breaks = c(1,10,30),
                        name = "Number of Attempts",
                        ) +
  
  # Add a transparent layer of contour plot
  stat_density2d_filled(alpha = 0.45, 
                        contour_var = "count",
                        breaks = c(0,0.035,0.5,1.5,3,6,10),
                        show.legend = FALSE) +
  scale_fill_manual(values = c("#FFFFFF","#fff6c5", "#ffb994", 
                               "#ff7b63", "#ff3e31", "#ff0000")) +
  scale_x_continuous(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0,60)) +
  # make x and y 1-to-1
  coord_equal() +
  # Add a Trae headshot pic on the top left corner
  annotation_custom(rasterGrob(trae_pic), xmin = 0, xmax = 20, 
                    ymin = 42, ymax = 62)

trae_hm
```

Both the width and length of the plots are **continuous** variables, the contour plot is also interpolated using the location and frequency of the data points to display the plays' unique favorite areas to attack.



# Top 9 players in MVP voting {.mybreak}

## Categorical Element (Facet)

There are tons of new faces in the **MVP ladder**, players like Luka Doncic, Ja Morant, and the Joker are all dominating the league in their own way. 

```{r, echo=FALSE, class.source="mybreak"}
# Data Cleaning
mvp_top5 <- pbox %>%
  # filter out the all-star game
  filter(!team_short_display_name %in% c("Team Durant","Team LeBron")) %>%
  #filter out the playeroffs 
  filter(season_type == 2) %>%
  filter(athlete_id %in% c("3112335","3059318","3032977",
                           "3136193","3945274","4065648",
                           "4279888","3975","2779")) %>%
  separate(fg, c("fgm","fga"), "-") %>%
  select(athlete_display_name, pts, fgm,fga, reb, ast, stl, blk, to, athlete_headshot_href) %>%
  group_by(athlete_display_name) %>%
  # Get the per-game statistics
  summarise(Point = sum(pts)/n(),
            Rebound = sum(reb)/n(),
            Assist = sum(ast)/n(),
            Steal = sum(stl)/n(),
            Block = sum(blk)/n(),
            #Turnover = sum(to)/n(),
            Percentage = sum(as.numeric(fgm)) / sum(as.numeric(fga)),
            pic = max(athlete_headshot_href)
  ) %>%
  
  # Sorting using order of MVP voting
  arrange(match(athlete_display_name, c("Nikola Jokic",
                                        "Joel Embiid",
                                        "Giannis Antetokounmpo",
                                        "Devin Booker",
                                        "Luka Doncic",
                                        "Jayson Tatum",
                                        "Ja Morant",
                                        "Stephen Curry",
                                        "Chris Paul"))) %>%
  remove_rownames %>% 
  column_to_rownames(var="athlete_display_name")

# Use the team color as the color of the radar plot, 
# the color in the dataset is not accurate
team_color <- c("FEC524", "006BB6", "003813",
                "E56020","00538C","007A33",
                "5D76A9","FFC72C","E56020")
max_min_mean <- data.frame(
  Point = c(max(mvp_top5$Point), 0,8), 
  Rebound = c(max(mvp_top5$Rebound),0, 4), 
  Assist = c(max(mvp_top5$Assist), 0,2), 
  Steal = c(max(mvp_top5$Steal), 0,0.5), 
  Block = c(max(mvp_top5$Block), 0,0.2), 
  #Turnover = c(0, max(mvp_top5$Turnover)), 
  Percentage = c(max(mvp_top5$Percentage), 0.35,0.45)
)

rownames(max_min_mean) <- c("Max", "Min", "Average")

mvp_top5_final <- rbind(max_min_mean, mvp_top5[,c(1:6)])

# Function to create a radarchart
create_radarchart <- function(data, color,
                              vlabels = colnames(data), 
                              vlcex = 0.7,
                              title = NULL, ...){
  radarchart(
    data,
    pcol = c(NA,color), 
    pfcol = c("#00000090",scales::alpha(color, 0.5)), 
    plwd = 2, plty = 1,
    cglcol = "grey", cglty = 1, cglwd = 0.8,
    axislabcol = "grey",
    vlcex = vlcex, vlabels = vlabels,
    title = title, ...
  )
}

# set up the layour for the 9 radar graphs
par(mar=c(1,1,1,1))
layout(matrix(c(1,2,5,8,1,3,6,9,1,4,7,10),ncol=3),
       heights=c(1,3,3,3))
plot.new()
text(0.5,0.5,"2021-22 MVP Ladder Radar Graphs",cex=2,font=2)

# Create the radar chart
for(i in 1:9){
  create_radarchart(
    data = mvp_top5_final[c(1,2,3, i+3), ],
    color = paste0("#", team_color[i]),
    title = paste0(i,". ",rownames(mvp_top5_final[i+3,]))
    )
}
```

The radar graphs are faceted by players, and the grey area in each plot is the average statistics for all NBA players.


# Score Plot for a NBA game

## Homebrewed

```{r, echo=FALSE, out.width="90%"}
# Homebrewed graph

# Data Cleaning
come_back <- pbp %>%
  filter(game_id == "401360535") %>%
  filter(scoring_play == TRUE) %>%
  mutate(scored_side = ifelse(team_id == home_team_id, "home", "away")) %>%
  select(home_score,away_score,scored_side) %>%
  arrange(home_score,away_score) %>%
  mutate(next_score_side = lead(scored_side,1)) %>%
  replace(is.na(.), "not_anything") %>%
  filter((scored_side != next_score_side)) %>%
  mutate(last_home_score = lag(home_score,1),
         last_away_score = lag(away_score,1)) %>%
  replace(is.na(.), 0)

# Set up the empty graph
par(mai = c(0.2,0.2,0.2,0.2))
plot(0, 0, type="n",
     xlim = c(0, 150),
     ylim = c(0, 100),
     axes=FALSE, xlab="",ylab="",main="", asp = 1)

# Draw triangles for each block of scoring
for (i in seq(nrow(come_back))){
  if(i%%2 == 1){
    polygon(x=come_back[i,c(5,1,6,5)]+14, y=c(75,75,25,75), col="#002B5C", border="black", lwd=1)
  } else {
    polygon(x=come_back[i,c(6,2,5,6)]+14, y=c(25,25,75,25), col="#C8102E", border="black", lwd=1)
  }
}

# indicate the winner
win_x = max(max(come_back$home_score),max(come_back$away_score))
lines(x = c(win_x+14.4,win_x+14.4), y = c(20,80), col="red", lty = 2, lwd=2)
if(come_back[nrow(come_back),3] == "away"){
  win_text_y = 25
} else {
  win_text_y = 75
}

# Add title/subtitle and labels
text(x=win_x+28, y=win_text_y+1, labels="Winner!!",col="red",cex = 1.5)
text(x=5, y=26, labels="Los Angeles \n Clippers",cex = 0.8)
text(x=5, y=75, labels="Washington \n Wizards", cex=0.8)
text(x=60, y=20, labels="Score")
text(x=46, y=78, labels="Clippers vs Wizards, Jan 25. 2022")
text(x=68, y=85, labels="Second Largest Comeback in NBA history",cex=1.4)
```

The graph above is created using R default **polygons** and **lines**. The horizontal axis represents the score for each team, with the bottom edge representing Clipper's score and the top edge representing Wizard's score. The lines between the top and bottom edges represent an instant of the game. The location where those lines connect with the top and bottom edge is the score for each team at that instant respectively. The polygons are colored with the teams' team color, the dark blue polygons represent a period of time when the Wizards are scoring, and the red polygons represent the period when the clippers are scoring. 

I didn't add any value on the edges because I believe that would add too many numbers and might confuse the audience, the purpose of this graph is to display which team was ahead at each period and the process of the game.

# Teams' Scoring Duo {.mybreak}

## Wildcard

The age of Big Threes is over now, more and more teams are aiming for only **2 core/stars** in their lineup. Lets check out the scoring duos of each team!

```{r myprettycode, echo=FALSE, fig.cap='Table', fig.height=6, out.width="80%", class.source="mybreak"}
# Data cleaning
best_duo <- pbox %>%
  # filter out the all-star game
  filter(!team_short_display_name %in% c("Team Durant","Team LeBron")) %>%
  #filter out the playeroffs 
  filter(season_type == 2) %>%
  separate(fg, c("fgm","fga"), "-") %>%
  group_by(athlete_display_name) %>%
  # Need an player to play at least 40 games
  filter(n() > 40) %>%
  summarise(Points.per.Game = sum(pts)/n(),
            Team_Name = max(team_name)) %>%
  group_by(Team_Name) %>%
  # Get the top 2 scorer in each team
  slice_max(order_by = Points.per.Game,n=2) %>%
  group_by(Team_Name) %>%
  # Identify the first and second in Duo
  summarise(Combined_ppg = round(sum(Points.per.Game),1),
            Player_1 = athlete_display_name[which.max(Points.per.Game)],
            Player_1_ppg = round(max(Points.per.Game),1),
            Player_2 = athlete_display_name[which.min(Points.per.Game)],
            Player_2_ppg = round(min(Points.per.Game),1)) %>%
  arrange(desc(Combined_ppg))

# Create the Huxtable
duo_tb <- hux(best_duo) %>%
  set_bold(1,everywhere) %>%
  set_bottom_border(1,everywhere, brdr(1, "double")) %>%
  set_right_border(everywhere, 1, brdr(1, "double")) %>%
  set_background_color(evens, everywhere, "grey95") %>%
  map_text_color(-1, c(2,4,6), 
                 by_colorspace("blue", "grey50", "red", colwise = TRUE)
                 )
font_size(duo_tb)  <-  15
height(duo_tb) <- 1
huxtable::width(duo_tb) <- 1
add_footnote(duo_tb, "*ppg - Points per Game")
```

The table is built using Huxtable, ordered by the **combined ppg** for each teams' scoring duo. The numeric columns are color coded, where red is the highest and blue is the lowest in the column. The ideal range of values in a table is 6-40, however, I believe in this case the number of value is acceptable since the table functions as a stats sheet. 






